\plaintitle

### 2016-09-21 drafts/labor/relationships.tex

\subsubsection{\pieceworkpers}
\citeauthor{clark1908cotton}, for instance,
relays his observations of textile mill pieceworkers and his realization that
``When he works by the day the Italian operative wishes to leave before the whistle blows,
but if he works by the piece he will work as many hours as it is possible for him to stand''

\citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
and \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
offering guidance on how best to manage pieceworkers
\cite{riisOtherSideLives,norton1900textile}.

\topic{Despite the intense growth of the piecework approach to renumeration,
this time was not without turmoil.}
As previously discussed, a number of worker organizations weighed in on
(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
\cite{american1921problem,richards1904anything}.
As mounting attention increasingly revealed problems in piecework as it related to workers,
workers themselves began to speak out about their frustration with this new regime.
\citeauthor{riisOtherSideLives}'s photo--documentary work
brought light to otherwise--invisible pieceworkers in their homes;
this and other events led to organizations representing
railway workers, mechanical engineers, and others
who began to mount advocacy in defense and favor of workers
\cite{american1921problem,richards1904anything,riisOtherSideLives}.
\citeauthor{10.2307/3827491} tells us of the match--girls strike of 1888,
one of the earliest and most famous successful worker strikes and perhaps
the beginning of ``militant trade unionism''
\cite{10.2307/3827491}.
It wasn't an exaggeration when \citeauthor{weyer1894history} later said that
``the match--girls' victory turned a new leaf in Trade Union annals'' \cite{weyer1894history};
in the 30 years since the match--girls strike,
the Trade Union Movement grew from
``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
\cite{webb1920history}.

\topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
--- albeit critical ---
case study in this space;
the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
to strike for
an individual minimum wage, among other demands
\cite{10.2307/2221944}.
\citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
\cite{10.2307/41829256}.
The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
\cite{jacoby1983union}.

\topic{The questions surrounding
the ways pieceworkers related to managers might be best answered by
the work that has been done in the emergence and proliferation of labor unions.}
The primary avenue for workers to interact with managers has been through
laborer advocacy groups such as the American Federation of Labor,
(one of the forerunners of the largest and
most politically influential labor union in the United States).
Looking through that lens, we find copious research on
the relationships between workers and requesters
\cite{levi2009union,ahlquist2013interest,
      mccallum2013global,jacoby1983union}.
One component of collectively negotiating with managers has been the process
of collective action, a topic which has been substantively explored
but is not quite yet answered
\cite{russell1982collective,olsonlogic}.

\topic{Answering how workers related to one another is arguably more challenging
for a number of reasons.}
For one thing, the research methods we typically associate with the exploratory study of cultures
--- Anthropology, and namely participant--observation, ethnography, etc\dots ---
didn't exist quite as we know them at the turn of the \nth{20} century, and
wouldn't for several more decades.
Still, we can look at primary sources, like \citetitle{american1921problem}
to give us some hint of how they related to each other
\cite{american1921problem}.

The driving force of American labor advocacy organizations was to get piecework railroad workers
to identify
``not only as railroad employees but also as members of the larger life of the community''
\cite{american1921problem}.
Doing this, 
\citeauthor{ostrom1990governing} and others argued,
would facilitate collective action and perhaps collective governance
\cite{ostrom1990governing,russell1982collective,olsonlogic}.
\citeauthor{riisOtherSideLives} had contributed to this sense of shared struggle and endurance
by the time \citetitle{american1921problem} was published
by documenting pieceworkers in their home--workplaces,
literally bringing to light the grim circumstances in which pieceworkers lived and worked
\cite{riisOtherSideLives}.

@@

\subsubsection{\whatchanged}
\topic{The differences between crowd workers and pieceworkers seem defined largely by
the differences in the places of work.}
Whereas it arguably became inevitable that workers would have a place to
meet, discuss, and collaborate
when they began sharing places of work,
online spaces make it much harder to do so.
Crowd workers can ``lurk'' and do tasks, or just do the occasional one--off task,
without any affiliation with
--- or even knowledge of ---
communities of peers
% \ali{multiple citations that labor unions came out of factories here}
\cite{miller2011understanding,mcinnis2016one,earl2011digitally}.

>>>>

\subsubsection{\pieceworkpers}
Early observers believed that workers were strongly motivated by the piecework model.
\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
``When he works by the day the Italian operative wishes to leave before the whistle blows,
but if he works by the piece he will work as many hours as it is possible for him to stand.''
Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.

Workers' relationships with employers quickly soured.
The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described, ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike, the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
% in the 30 years since the match--girls strike,
% the Trade Union Movement grew from
% ``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
% \cite{webb1920history}.
This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
\cite{jacoby1983union}.

Soon, many worker organizations were weighing in on (or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time~\cite{american1921problem,richards1904anything}.
As mounting attention increasingly revealed problems in piecework's treatment of workers, theworkers themselves began to speak out about their frustration with this new regime.
Organizations representing railway workers, mechanical engineers, and others began to mount advocacy in defense of workers
\cite{american1921problem,richards1904anything}.

% \topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
% The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
% --- albeit critical ---
% case study in this space;
% the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
% to strike for
% an individual minimum wage, among other demands
% \cite{10.2307/2221944}.
% \citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
% \cite{10.2307/41829256}.
% The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
% \cite{jacoby1983union}.

Pieceworkers' relationships with their employers eventually developed a pattern of using 
% \topic{The questions surrounding
% the ways pieceworkers related to managers might be best answered by
% the work that has been done in the emergence and proliferation of labor unions.}
% The primary avenue for workers to interact with managers has been through
laborer advocacy groups~\cite{levi2009union,ahlquist2013interest,
      mccallum2013global,jacoby1983union}.
% (one of the forerunners of the largest and
% most politically influential labor union in the United States).
% Looking through that lens, we find copious research on
% the relationships between workers and requesters
% \cite{levi2009union,ahlquist2013interest,
%       mccallum2013global,jacoby1983union}.
Collective action grew to become a central component of negotiating with managers~\cite{russell1982collective,olsonlogic}.

\topic{Less is known about how pieceworkers related to each other.}
For one thing, the research methods we typically associate with the exploratory study of cultures
--- Anthropology, and namely participant--observation, ethnography, etc\dots ---
didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
% Still, we can look at primary sources, like \citetitle{american1921problem}
% to give us some hint of how they related to each other
% \cite{american1921problem}.
Primary sources indicate that labor organizations wished for workers to identify as a collective group, 
% The driving force of American labor advocacy organizations was to get piecework railroad workers
% to identify
``not only as railroad employees but also as members of the larger life of the community''
\cite{american1921problem}.
Doing this, 
\citeauthor{ostrom1990governing} and others argued,
would facilitate collective action and perhaps collective governance
\cite{ostrom1990governing,russell1982collective,olsonlogic}.
\citeauthor{riisOtherSideLives} also contributed to this sense of shared struggle and endurance
% by the time \citetitle{american1921problem} was published
by documenting pieceworkers in their home--workplaces,
literally bringing to light the grim circumstances in which pieceworkers lived and worked
\cite{riisOtherSideLives}.

@@

\subsubsection{\whatchanged}

\begin{comment}
\end{comment}

\topic{The differences between crowd workers and pieceworkers seem defined largely by
the differences in the places of work.}
Whereas it arguably became inevitable that workers would have a place to
meet, discuss, and collaborate
when they began sharing places of work,
online spaces make it much harder to do so.
Crowd workers can ``lurk'' and do tasks, or just do the occasional one--off task,
without any affiliation with
--- or even knowledge of ---
communities of peers
% \ali{multiple citations that labor unions came out of factories here}
\cite{miller2011understanding,mcinnis2016one,earl2011digitally}.
